Teaching Debugging Collaboratively---Midway Report

### 2020-05-08 paper/midway.tex
Section 2 done!

As Michaeli and Romeike point out, there is a lack of research into
the value of teaching debugging in higher education.

Research into how to best teach debugging is self-admittedly sparse.
Chan et al. allow that ``in general research on how to improve
debugging is sporadic''---an observation that leads them to research a
framework to reduce the complexity of teaching
debugging\cite{10.1145/3286960.3286970}.  To organize their framework,
they split debugging knowledge into 5 categories: Domain, System,
Procedural, Strategic, and Experiential.  They then review different
debugging tools and teaching aids---from those that involve writing
code to games---and map tools to the knowledge areas they seek to
address.  After an evaluation of a host of different tools, they claim
a few significant faults in current debugging teaching platforms.  The
two of which I seek to address follow:
\begin{itemize}
\item A lack of tools addressing system knowledge (an understanding of
  the program to be debugged).
\item A lack of back-tracing ability/coverage.
\end{itemize}

>>>>

As Michaeli and Romeike point out, there is a lack of research into
the value of teaching debugging in higher education.  None of the
research they found placed much focus on explicitly teaching
debugging.  Chmiel and Loui studied whether students who were provided
with debugging tools and frameworks performed better on tests or spent
less time on assignments than those who were
not\cite{10.1145/971300.971310}.  They weren't able to find conclusive
evidence towards better performance on tests or assignments, though
they did find that students in the treatment group felt more confident
in their debugging abilities.  Unfortunately their study didn't
involve extended explicit teaching of debugging--use of the tools was
voluntary, and variations in the students' individual abilities made
the data difficult to evaluate.\par

Despite the lack of higher education-specific research, I feel that
the value of teaching debugging has still been demonstrated.  The
research discussed all finds that K12 and college students alike
commonly resort to sporadic debugging techniques when beginning to
learn.  Given this, I see no reason why the conclusive and positive
research toward the benefit of explicitly teaching debugging to K12
students should not apply to their collegiate counterparts.\par

Research into how to best teach debugging is also self-admittedly
sparse.  Chan et al. allow that ``in general research on how to
improve debugging is sporadic''---an observation that leads them to
research a framework to reduce the complexity of teaching
debugging\cite{10.1145/3286960.3286970}.  To organize their framework,
they split debugging knowledge into 5 categories: Domain, System,
Procedural, Strategic, and Experiential.  They then review different
debugging tools and teaching aids---from those that involve writing
code to games---and map tools to the knowledge areas they seek to
address.  After an evaluation of a host of different tools, they claim
to find a few significant faults in current debugging teaching
platforms.  The two of which I seek to address follow:

\begin{enumerate}
\item A lack of back-tracing ability/coverage.
\item A lack of tools addressing system knowledge (an understanding of
  the program to be debugged).
\end{enumerate}

Basing my collaborative debugging platform on RR solves the problem of
back-tracing ability and coverage.  RR is a ``time traveling''
addition to GDB, that allows for unlimited recording and replaying
code execution deterministically, both backwards and
forwards\cite{DBLP:journals/corr/OCallahanJFHNP17}.  By integrating
the debugger with tools to visualize the execution and code space in
an accessible way, I hope to address the issue of system knowledge.
Though the system knowledge gained may be primarily applicable to low
level programming, skills such as visualizing code execution and
memory space are applicable to all types of coding.  These solutions
are discussed in depth in the following sections.
